Adaptive Sequential Model Selection

### 2015-07-11 paper.tex
rob made small changes

\begin{abstract}
  Many model selection algorithms produce a ``path'' of fits specifying a sequence of increasingly complex models. Given such a sequence and the data used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data. Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path, accounting for the fact that the model path is determined adaptively using the data. In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
  We also show that the selected models test  bears a close relationship to parametric bootstrap sampling,

@@

  Then, $\hk$ controls $\Err$ at $\alpha$
  whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
\end{proposition}
\begin{proof}
  Fix any values $k_0\in \{0,\ldots,d, \infty\}$ and $q_{1:k_0} \in [0,1]^{k_0}$. We can construct a problem with $\tilde p_k \sim \delta_{q_k}$ for $k\leq k_0$ and  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$. 
  \RTcomment{Will; what are the qk?  you mean we just choose and fix some values qk? Say so.} \WFcomment{Is it clearer now?}

@@

\begin{table}
\begin{center}
\begin{tabular}{ l | r|r| r}
Procedure& Complete Null & Incremental Null\\
\hline
LAR /lasso& Covariance test/Saturated&Spacing test/Saturated \\
        & Next-$\lambda$ test/Selective& \\
        \hline
Forward Stepwise &maxT, maxT-identify/Selective &Gaussian pivot/Saturated\\
\hline
\end{tabular}
\end{center}
\caption{\em Summary of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
Tests are described by the null hypothesis that they test (complete or incremental)  and the underlying model (saturated or selective).
The maxT, maxT-identify and next-$\lambda$ tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
pivot tests are given in \citet{spacings}. The latter paper shows that the covariance and spacing test are asymptotically equivalent for the first null step, when the complete null holds,}
\label{tab:testsummary}
\end{table}
Figure  \ref{fig:larexample} shows an example of the LAR/lasso tests with $N=50, p=10$ and two strong signals.
We plot the p-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
We see that the next-$\lambda$ test is more powerful , especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
while the covariance test becomes conservative in Step 4, as we move further into the null part of the sequence.
\begin{figure}
  \includegraphics[width=\textwidth]{figs/larexample.pdf}
  \caption[fig:larexample]{\em  Cumulative distribution functions of p-values for LAR/lasso tests from simulations with $N=50,  p=10$ and two strong signals.}
  \label{fig:larexample}
\end{figure}

@@

Note that  (\ref{eq:boot})  has the form of  (conditional) parametric bootstrap sampling, in which the bootstrap residuals projected into the subspace
orthogonal to the current fit. In similar fashion, one might use a instead a nonparametric version of (\re{eq:boot}), sampling
with replacement from the model residuals and then projecting them orthogonal to the current model.

Figure \ref{fig:comparison} shows and example with $n=100,p=20$ and one strong signal, computing the max-t pvalue at Step 3.
The ``exact'' answer  was determined from bootstrap sampling with $20,000$ replications.
The Figures show the bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
over 10 different starting seeds.
\begin{figure}[htp]
\centering
  \includegraphics[width=\textwidth]{figs/comparison.pdf}
  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-t pvalue at Step 3.  Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
over 10 different starting seeds.
}
  \label{fig:comparison}
\end{figure}
We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for  8000 samples.
Hence we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low to make is practical.

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
  Method, Stopping rule& $p_{\text screen}$ & FWER$_{\text mod}$ & FDR$_{\text mod}$ &FDR$_{\text var}$ \\
% & \$p\_\{$\backslash$text\{screen\}\} & \$$\backslash$text\{FWER\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{variable\}\}\$ \\ 
  \hline
Nominal, Simple & 0.832 & 0.252 & 0.097 & 0.171 \\ 
  Nominal, Forward & 0.931 & 0.896 & 0.345 & 0.426 \\ 
  Nominal, Strong & 0.003 & 0.000 & 0.000 & 0.035 \\ 
  \hline
  MaxT, Simple & 0.369 & 0.000 & 0.002 & 0.042 \\ 
  MaxT, Forward & 0.668 & 0.038 & 0.026 & 0.080 \\ 
  MaxT, Strong & 0.025 & 0.006 & 0.004 & 0.038 \\ 
  \hline
  MaxT-Identify, Simple & 0.404 & 0.000 & 0.002 & 0.042 \\ 
  MaxT-Identify, Forward & 0.688 & 0.039 & 0.026 & 0.082 \\ 
  MaxT-Identify, Strong & 0.024 & 0.003 & 0.002 & 0.037 \\ 
  \hline
  Saturated, Simple & 0.000 & 0.000 & 0.000 & 0.029 \\ 
  Saturated, Forward & 0.013 & 0.001 & 0.000 & 0.030 \\ 
  Saturated, Strong & 0.000 & 0.000 & 0.000 & 0.033 \\ 
  \hline
  Knockoffs & 0.252 &  &  & 0.103 \\ 
   \hline
\end{tabular}
\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.}
\label{tab:stopping}
\end{table}

>>>>

\begin{abstract}
  Many model selection algorithms produce a ``path'' of fits specifying a sequence of increasingly complex models. Given such a sequence and the data used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data. Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path, accounting for the fact that the model path is determined adaptively using the data. In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
  We also show that the selected models test  bears a close relationship to parametric bootstrap sampling.

@@

  Then, $\hk$ controls $\Err$ at $\alpha$
  whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
\end{proposition}
\begin{proof}
  Fix any values $k_0\in \{0,\ldots,d, \infty\}$ and $q_{1:k_0} \in [0,1]^{k_0}$. We can construct a problem with $\tilde p_k \sim \delta_{q_k}$ for $k\leq k_0$ and  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$.

@@

\begin{table}
\begin{center}
\begin{tabular}{ | l | r|r| r}
        \hline
Procedure& Complete Null & Incremental Null\\
\hline
LAR/lasso& Covariance test/Saturated&Spacing test/Saturated \\
        & Next-$\lambda$ test/Selective& \\
        \hline
Forward Stepwise &maxT, maxT-identify/Selective &Gaussian pivot/Saturated\\
\hline
\end{tabular}
\end{center}
\caption{\em Summary of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
Tests are described by the null hypothesis that they test (complete or incremental)  and the underlying model (saturated or selective).
The maxT, maxT-identify and next-$\lambda$ tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
pivot tests are given in \citet{spacings}. The latter paper shows that the covariance and spacing test are asymptotically equivalent for the first null step, when the complete null holds,}
\label{tab:testsummary}
\end{table}
Figure  \ref{fig:larexample} shows an example of the LAR/lasso tests with $N=50, p=10$ and two strong signals.
We plot the p-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
We see that the next-$\lambda$ test is the most powerful , especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
while the covariance test becomes conservative in step 4, as we move further into the null part of the sequence.
\begin{figure}
  \includegraphics[width=\textwidth]{figs/larexample.pdf}
  \caption[fig:larexample]{\em  Cumulative distribution functions of p-values for LAR/lasso tests from simulations with $N=50,  p=10$ and two strong signals.}
  \label{fig:larexample}
\end{figure}

@@

Note that  (\ref{eq:boot})  has the form of  (conditional) parametric bootstrap sampling, in which the bootstrap residuals projected into the subspace
orthogonal to the current fit.  The typical justification for the bootstrap is asymptotic, relying on the convergence of the parameter estimates to the true parameters.
But something magical  occurs in the selective sampling scheme (\ref{eq:boot}): this  sampling fixes the value of $ X_{E_{k-1}}^Ty$ and this conditioning
removes the dependence on nuisance parameters.
Hence  (\ref{eq:boot}) yields the exact (non-asymptotic) conditional distributions and p-values.

Figure \ref{fig:comparison} shows an example with $n=100, p=20$ and one strong signal, computing the max-t pvalue at Step 3.
The ``exact'' answer  was determined from bootstrap sampling with $20,000$ replications.
The Figures show the bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
over 10 different starting seeds.
\begin{figure}[htp]
\centering
  \includegraphics[width=\textwidth]{figs/comparison.pdf}
  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-t pvalue at Step 3.  Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
over 10 different starting seeds.
}
  \label{fig:comparison}
\end{figure}
We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for  8000 samples.
Hence we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low to make is practical.

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
  Method, Stopping rule& $p_{\text screen}$ & FWER$_{\text mod}$ & FDR$_{\text mod}$ &FDR$_{\text var}$ \\
% & \$p\_\{$\backslash$text\{screen\}\} & \$$\backslash$text\{FWER\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{variable\}\}\$ \\ 
  \hline
Nominal, Simple & 0.832 & 0.252 & 0.097 & 0.171 \\ 
  Nominal, Forward & 0.931 & 0.896 & 0.345 & 0.426 \\ 
  Nominal, Strong & 0.003 & 0.000 & 0.000 & 0.035 \\ 
  \hline
  MaxT, Simple & 0.369 & 0.000 & 0.002 & 0.042 \\ 
  MaxT, Forward & 0.668 & 0.038 & 0.026 & 0.080 \\ 
  MaxT, Strong & 0.025 & 0.006 & 0.004 & 0.038 \\ 
  \hline
  MaxT-Identify, Simple & 0.404 & 0.000 & 0.002 & 0.042 \\ 
  MaxT-Identify, Forward & 0.688 & 0.039 & 0.026 & 0.082 \\ 
  MaxT-Identify, Strong & 0.024 & 0.003 & 0.002 & 0.037 \\ 
  \hline
  Saturated, Simple & 0.000 & 0.000 & 0.000 & 0.029 \\ 
  Saturated, Forward & 0.013 & 0.001 & 0.000 & 0.030 \\ 
  Saturated, Strong & 0.000 & 0.000 & 0.000 & 0.033 \\ 
  \hline
  Knockoff & 0.252 &  &  & 0.103\\ 
  Knockoff+& ??   &  &    &   0.000? \\
   \hline
\end{tabular}
\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.}
\label{tab:stopping}
\end{table}
